\abstract

% This work explores the application of the phase-field method to the simulation of fluid-driven fracture and proposes a framework to make simulations faster and more robust. It begins by addressing a modeling challenge related to the application of pressure loads on the diffuse crack interfaces. During the verification process, a new J-Integral for pressurized fractures in the phase-field context is developed.

% Then, the focus turns into a hybrid method to describe fracture propagation. The so-called multi-resolution method uses a combination of enrichment schemes with the phase-field method to address the complex fluid-fracture interaction that happens in hydraulic fracture. On one hand, the phase-field method alleviate some of the difficulties associated with the geometric evolution of the fracture, which are usually the limiting aspect of purely enrichment-based schemes. On the other hand, the discrete representation allows for a better treatment of the fluid loads and crack apertures, which are the main challenges associated with phase-field approaches. 

% The multi-resolution method is first presented in a simplified scheme to treat two-dimensional problems, which is used to verify the framework against well-known analytical solutions. It is then extended to three-dimensions. A robust algorithm to handle planar cracks in 3D is developed and its extension to nonplanar cases is discussed. Opportunites for improvements and extensions are then discussed, paving the road for future work in the topic.

This work advances a phase-field method for fluid-driven fractures and proposes a robust and efficient discretization framework. It begins by addressing a modeling challenge related to the application of pressure loads on diffuse crack surfaces. Along the way, a new J-Integral for pressurized fractures in a regularized context is developed. 

Then, the focus turns to a hybrid method to model fluid-driven fracture propagation. A so-called multi-resolution method is presented that uses a combination of enrichment schemes with the phase-field method to address the complex fluid-fracture interaction that occurs during hydraulic fracture. On one hand, the phase-field method alleviates some of the difficulties associated with the geometric evolution of the fracture, which are usually the limiting aspect of purely enrichment-based schemes. On the other hand, the discrete representation 
allows for a better treatment of the fluid loads and crack apertures, which are the main challenges associated with phase-field approaches. 

The multi-resolution method is first presented in a simplified scheme to treat two-dimensional problems.  Various benchmark problems are used to verify the framework against well-known analytical solutions. The method is then extended to three-dimensions. A robust algorithm to handle planar cracks in 3D is developed and its extension to non-planar cases is discussed. Finally, opportunities for improvements and extensions are discussed, paving the road for future work in this area.